---
title: "ML_Assignment1"
date: "2023-01-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Loading library

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
library(lmtest)
library(knitr)
library(kableExtra)

set.seed(1)
```

## Question 1:
#### Construct a table providing summaries of the quantitative features of the dataset.

```{r}
bcdata = read_csv(file = "./data/Assignment1_bcdata.csv")
bcdata = janitor::clean_names(bcdata)
```
```{r}
summarized_df = bcdata %>% rename (mcp1 = mcp_1) %>%
  select(-c(classification)) %>% 
  summarise(across(everything(), list(mean=mean, median=median, min = min, max = max)))  %>% 
   pivot_longer(age_mean:mcp1_max,
                 names_to = "Type", 
                 values_to = "Value") %>% 
  separate(Type, c('Quantitative features', 'Summaries')) %>% 
  pivot_wider (names_from = "Summaries", 
  values_from = "Value") %>% 
knitr::kable (caption = "Summaries of the Quantitative Features", align = c ("c", "c"), digits = 2) %>%  kable_paper("striped", full_width = F) %>% 
  column_spec(1, bold = T)


summarized_df
  
```
*Comments: The table is summaries of the quantitative features. And their units are as follows: Age (years), BMI (kg/m2), Glucose (mg/dL), Insulin (μU/mL), HOMA-IR (), Leptin (ng/mL),  Adiponectin (μg/mL),  Resistin (ng/mL),  MCP-1(pg/dL).*


## Question 2:
#### Recode BMI into the WHO-defined categories below

```{r}
Q2_df = bcdata %>% mutate(bmi_group = 
                    case_when(bmi < 16.5 ~ 'Severely underweight',
                              bmi < 18.5 ~ "Underweight",
                              bmi <= 24.9 ~ "Normal weight",
                              bmi <= 29.9 ~"Overweight",
                              bmi <= 34.9 ~"Obesity class I",
                              bmi <= 39.9 ~"Obesity class II",
                              bmi >= 40~ "Obesity class III"))
  Q2_df         
```
*Comments: The continuous variable `BMI` was transferred into a categorical variable `bmi_group` based on the WHO-defined categories. And there were only 5 categories*


## Question 3:
#### Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category.

```{r}
Q3_cases = Q2_df %>% group_by(bmi_group) %>% 
  filter(classification==2) %>% 
  summarize(count = n()) %>% 
  mutate(cases_no = count) %>% select(!count)

Q3_controls = Q2_df %>% group_by(bmi_group) %>% 
  filter(classification==1) %>% 
  summarize(count = n())  %>% 
  mutate(controls_no = count) %>% select(!count)
```

```{r}
Q3_total=full_join(Q3_cases,Q3_controls, by=c("bmi_group")) 
 
Q3_total[is.na(Q3_total)] = 0

Q3_total = Q3_total %>% mutate (total_no =cases_no + controls_no ) %>% 
  mutate (cases_percent= (cases_no / total_no)*100,
          controls_percent = (controls_no / total_no)*100) %>% 
  select(bmi_group, cases_percent, controls_percent) %>% 
  mutate(bmi_group = fct_relevel(bmi_group, c("Underweight", "Normal weight",
                       "Overweight","Obesity class I","Obesity class II"))) %>% 
   arrange(bmi_group) %>% 
  pivot_longer(cases_percent:controls_percent,
                 names_to= "Type",
                 values_to = "Percent")

```

```{r}

bar_chart = Q3_total %>% ggplot(aes(x=bmi_group,y=Percent,fill=Type)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_brewer(palette = "Pastel2")+
 geom_text(aes(label=sprintf("%0.2f", Percent)), position=position_dodge(width=0.9), vjust=-0.25, size=3) +
  labs(title = "The proportion of breast cancer cases and controls within each BMI category", x = "BMI category", y = "The type of participants")

bar_chart
```

## Question 4:
#### Construct a logistic regression model using breast cancer classification as the outcome and glucose, HOMA-IR, leptin, BMI (continuous) and age as the independent variables.

```{r}
bcdata_df = bcdata %>% mutate(classification = recode_factor(classification,
                                                             "1"="0",
                                                             "2"="1"))
fit_logistic = 
  glm(classification ~ glucose + homa + leptin + bmi + age, family=binomial(link="logit"), data = bcdata_df)
```

```{r}
fit_logistic %>% 
  broom::tidy() %>% 
   mutate(Lower_CI = (estimate -1.96*std.error),
         Upper_CI = (estimate +1.96*std.error)) %>% 
  select(term, estimate,std.error, Lower_CI, Upper_CI, statistic, p.value) %>% 
  kbl(caption     = "Effect of Selected Predictors on breat cancer classification"
    , col.names   = c("Predictors", "Beta Estimate","SE", "Lower bound of 95% CI","Upper bound of 95% CI", "t-statistic", "p-value"),
    digits= 2) %>%
 kable_paper("striped", full_width = F) %>% 
  column_spec(1, bold = T)
 
```
*Comments: With all the other variables held constant, if 1-unit changes in HOMA-IR, the beta estimate is 0.27 and its related 95% confidence interval from -0.06 to 0.61*


## Question 5:
#### Construct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables.

```{r}
fit_linear = lm (insulin ~ bmi + age + glucose, data = bcdata)
```

```{r}
fit_linear %>% 
   summary() %>%
   broom::tidy() %>% 
  mutate(Lower_CI = (estimate -1.96*std.error),
         Upper_CI = (estimate +1.96*std.error)) %>% 
  select(term, estimate, std.error, Lower_CI, Upper_CI, statistic, p.value) %>%
  kbl(caption= "Effect of Selected Predictors on insulin level"
    , col.names   = c("Predictor", "Beta Estimate", "SE","Lower bound of 95% CI",
                      "Upper bound of 95% CI","t-statistic", "p-value"),
     digits= 6) %>%
  kable_paper("striped", full_width = F) %>%
  column_spec(1, bold = T)
  
```
*Comments: With all the other variables held constant, if 1-unit changes in age, the beta estimate is -0.054 and its related 95% confidence interval from -0.156 to 0.048.*
