---
title: "ML_Assignment1"
date: "2023-01-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Loading library

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
```

## Question 1:Construct a table providing summaries of the quantitative features of the dataset.

```{r}
bcdata = read_csv(file = "./data/Assignment1_bcdata.csv")
bcdata = janitor::clean_names(bcdata)
```
```{r}
summarized_df = bcdata %>% rename (mcp1 = mcp_1) %>%
  select(-c(classification)) %>% 
  summarise(across(everything(), list(mean=mean, median=median, min = min, max = max)))  %>% 
   pivot_longer(age_mean:mcp1_max,
                 names_to = "Type", 
                 values_to = "Value") %>% 
  separate(Type, c('Quantitative features', 'Summaries')) %>% 
  pivot_wider (names_from = "Summaries", 
  values_from = "Value") %>% 
  
knitr::kable (caption = "Summaries of the Quantitative Features", align = c ("c", "c"), digits = 2)

summarized_df
  
```
*Comments: The table is summaries of the quantitative features. And their units are as follows: Age (years), BMI (kg/m2), Glucose (mg/dL), Insulin (μU/mL), HOMA-IR (), Leptin (ng/mL),  Adiponectin (μg/mL),  Resistin (ng/mL),  MCP-1(pg/dL).*


## Question 2:Recode BMI into the WHO-defined categories below

```{r}
Q2_df = bcdata %>% mutate(bmi_group = 
                    case_when(bmi < 16.5 ~ 'Severely underweight',
                              bmi < 18.5 ~ "Underweight",
                              bmi <= 24.9 ~ "Normal weight",
                              bmi <= 29.9 ~"Overweight",
                              bmi <= 34.9 ~"Obesity class I",
                              bmi <= 39.9 ~"Obesity class II",
                              bmi >= 40~ "Obesity class III"))
  Q2_df         
```
*Comments: The continuous variable `BMI` was transferred into a categorical variable `bmi_group` based on the WHO-defined categories.*


## Question 3:Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category.

```{r}
Q3_cases = Q2_df %>% group_by(bmi_group) %>% 
  filter(classification==2) %>% 
  summarize(count = n()) %>% 
  mutate(cases_no = count) %>% select(!count)

Q3_controls = Q2_df %>% group_by(bmi_group) %>% 
  filter(classification==1) %>% 
  summarize(count = n())  %>% 
  mutate(controls_no = count) %>% select(!count)
```

```{r}
Q3_total=full_join(Q3_cases,Q3_controls, by=c("bmi_group")) 
 
Q3_total[is.na(Q3_total)] = 0

Q3_total = Q3_total %>% mutate (total_no =cases_no + controls_no ) %>% 
  mutate (cases_percent= (cases_no / total_no)*100,
          controls_percent = (controls_no / total_no)*100)
```

```{r}
bar_plot= 
```

